---
title: "Evaluation of HVAC and COVID-19 Growth Rates"
author: "Bill Kayser"
date: "August 25, 2020"
echo: TRUE
---

```{r results='hide', message=F, include=F}
library(survey)
library(tidyverse)
# library(scales)
library(knitr)
theme_set(theme_light())
# source('../utils/utils.R')
knitr::opts_chunk$set(echo=F)
```


<style type="text/css">
.table {
    width: inherit;
    max-width: 100%;
    margin-bottom: 20px;
}
.math.display {
  font-size: 28px;
}
</style>


# Introduction

## Load 2015 RECS Survey Data

The Residential Energy Consumption Surveys (RECS) are used to estimate energy usage in homes.  They are conducted by the US Energy Information Administration.  Data and information can be downloaded from [https://www.eia.gov/consumption/residential/data/2015/index.php?view=microdata](https://www.eia.gov/consumption/residential/data/2015/index.php?view=microdata).

The survey data uses replicate weights using Fay's method:
   
   Dippo, Cathryn S., Robert E. Fay, and David H. Morganstein. 1984. “Computing Variances
   From Complex Samples With Replicate Weights.” In Proceedings of the Survey Research
   Methods Section. Washington, DC: American Statistical Association: 489–494.


```{r}

recs <- read.csv('../data/recs2015_public_v4.csv') %>%
  select(DOEID,  # record UID
         REGIONC, DIVISION,   # Census region and division
         NHSLDMEM, # Number of household members
         CDD30YR,  # 30 Year average, cooling days 65 degrees,
         HDD30YR,  # 30 Year average, heating days 65 degrees
         KWHCOL,   # Electricity usage for cooling, kilowatt hours) 
         NWEIGHT,  # total sample weight
         starts_with('BRRWT')) %>%  # BRR weights
         mutate(REGIONC = as.factor(REGIONC),
                DIVISION = as.factor(DIVISION))
```

The weights are used to calculate Relative Standard Error as a percentage of the mean value.
The code for calculating these RSEs is detailed in the [reference document](https://www.eia.gov/consumption/residential/data/2015/pdf/microdata_v3.pdf) on the EIA page above.

```{r}
sampweights <- recs$NWEIGHT
brrwts <- select(recs, starts_with("BRRWT"))
des <- svrepdesign(weights=sampweights, 
                   repweights=brrwts, 
                   type="Fay", 
                   rho=0.5, 
                   mse=TRUE,
                   data=recs)
des
```


I stopped here because breaking down this data by census division is not going to show anything informative, even if there is an association.  Too many other candidate dependencies, such as political influence, varying mitigation strategies, social patterns.  In otherwords, the hot regions tend to lean republican, which tend to lean to less restrictive measures.  

Might be different if you could do it state by state.


