---
title: "Cleaning COVID-19 Data from Johns Hopkins"
author: "Bill Kayser"
output:
  html_notebook: 
    toc: yes
  html_document: default
  pdf_document: 
    toc: yes
echo: TRUE
---

```{r results='hide', message=F, include=F}
library(tidyverse)
library(scales)
library(caTools)
source('../utils/utils.R')
knitr::opts_chunk$set(echo=F)
#old_theme <- theme_set(ggthemes::theme_fivethirtyeight())
old_theme <- theme_set(ggthemes::theme_few())


```


<style type="text/css">
.table {
    width: inherit;
    max-width: 100%;
    margin-bottom: 20px;
}
.math.display {
  font-size: 28px;
}
</style>

# Load Datasets

```{r echo=T}
cvdata.us <- readRDS('../data/cvdata.us.RDS')
cvdata.us.by_state <- readRDS('../data/cvdata.us.by_state.RDS')
cvdata.i18n <- readRDS('../data/cvdata.i18n.RDS')
```

# Data Exploration

For the US Data I'm just showing a few states so the plots are more readable.  This is just preliminary exploration as I get to know the data and clean it for other purposes.

Let's first set up some parameters for the graph's X axis.

```{r}
start_date <- mdy('02/26/2020')
end_date <- max(cvdata.us$Date)
date.breaks <- seq(end_date,
              start_date,
              by="-1 week") %>% rev()
if (date.breaks[1] != start_date) {
  date.breaks <- c(date.breaks[1]-days(7), date.breaks)
  start_date <- date.breaks[1]
}
```


### Top Countries infected

```{r top_countries, fig.width=10}
countries <- 
  filter(cvdata.i18n, Date == end_date) %>%
  arrange(desc(Cases)) %>% 
  head(10)

ggplot(countries) +
  aes(fill=Country,
      x=Country,
      y=Cases) +
  geom_col(show.legend=F) +
  scale_y_continuous(labels=comma) +
  xlab(NULL)
rm(countries)
```
```{r fig.width=5, fig.height=4}
cvdata.i18n %>%
  arrange(desc(Date)) %>%
  group_by(Key) %>%
  summarize(Deaths.Per100K = first(Deaths.Per100K)) %>%
  arrange(desc(Deaths.Per100K)) %>%
  head(50) %>%
  ggplot() +
  aes(x=reorder(Key, Deaths.Per100K), y=Deaths.Per100K, fill=Deaths.Per100K) +
  geom_col() +
  scale_fill_gradient(low='#00FFAA', high='red', trans='log10') +
  geom_text(aes(label=round(Deaths.Per100K)), size=4, nudge_y=2) +
  coord_flip() +  
  xlab(NULL) +
  ylab('Deaths per 100K Population') +
  ggtitle('Countries Ranked by Per Capita Mortality Rate') +
  theme(line = element_blank(), 
        panel.border = element_blank(),
        axis.text.x = element_blank(),
        legend.position=c(0.5,0.5))
```

```{r fig.width=5, fig.height=4}
cvdata.i18n %>%
  arrange(desc(Date)) %>%
  group_by(Key) %>%
  summarize(Cases.Per100K = first(Cases.Per100K)) %>%
  arrange(desc(Cases.Per100K)) %>%
  head(50) %>%
  ggplot() +
  aes(x=reorder(Key, Cases.Per100K), y=Cases.Per100K, fill=Cases.Per100K) +
  geom_col() +
  scale_fill_gradient(low='#00FFAA', high='red', trans='log10') +
  geom_text(aes(label=round(Cases.Per100K)), size=4, nudge_y=45) +
  coord_flip() +  
  xlab(NULL) +
  ylab('Cases per 100K Population') +
  ggtitle('Countries Ranked by Per Capita Infection Rate') +
  theme(line = element_blank(), 
        panel.border = element_blank(),
        axis.text.x = element_blank(),
        legend.position=c(0.5,0.5))
```


```{r fig.width=5, fig.height=6}

cvdata.i18n %>%
  arrange(desc(Date), desc(Cases)) %>%
  head(50) %>%
  ggplot() +
  aes(x=reorder(Key, Cases), y=Cases, fill=Cases) +
  geom_col() +
  scale_fill_gradient(low='#00FFAA', high='red', trans='log10') +
  geom_text(aes(label=comma(signif(Cases, 3))), size=4, nudge_y=80000) +
  coord_flip() +  
  xlab(NULL) +
  ylab('Confirmed Cases') +
  ggtitle('Countries Ranked by Confirmed Cases') +
  theme(line = element_blank(), 
        panel.border = element_blank(),
        axis.text.x = element_blank(),
        legend.position=c(0.5,0.5))

```



```{r fig.width=5, fig.height=6}

cvdata.i18n %>%
  filter(Date == end_date, Cases.Growth5 > 0) %>%
  arrange(desc(Cases.Growth5)) %>%
  head(50) %>%
  ggplot() +
  aes(x=reorder(Key, Cases.Growth5), y=Cases.Growth5, fill=Cases.Growth5) +
  geom_col(show.legend=F) +
  scale_fill_gradient(low='#00FFAA', high='red', trans='log10') +
  geom_text(aes(label=percent(round(Cases.Growth5, 3))), size=4, nudge_y=0.006) +
  coord_flip() +  
  xlab(NULL) +
  ylab('Growth in Confirmed Cases') +
  ggtitle('Countries Ranked by Rate of Growth') +
  theme(line = element_blank(), 
        panel.border = element_blank(),
        axis.text.x = element_blank())
  #scale_y_log10()
```

### International Summary

```{r}
as_of <- end_date
totals <- filter(cvdata.i18n, Date == as_of) %>%
  select(Cases, Deaths, Recovered, Population) %>%
  colSums()
```
As of `r format(as_of, "%A, %B %e, %Y")`, the total number of confirmed cases worldwide is `r comma(signif(totals['Cases'], 3))`.

The total number of deaths is `r comma(signif(totals['Deaths'], 3))`.

The number of reported recoveries is `r comma(signif(totals['Recovered'], 3))`.

The overall mortality rate based on reported deaths and confirmed cases is `r percent(signif(totals['Deaths']/totals['Cases'], 2))`.


### International Cases

```{r fig.width=6}
d0 <- filter(cvdata.i18n, Deaths >= 5 & Country %in% c("China, Hubei", "Spain", "France", "Germany","Iran", "Italy", "Korea, South", "United Kingdom", "US")) %>%
  mutate(Country = as.character(Country)) %>%
  filter(!is.na(Population))

d1 <- split(d0, d0$Country)
d2 <- lapply(d1, function(table) {
  table$Day <- 1:nrow(table)
  table
})

labels <- lapply(d2, tail, 1) %>% bind_rows()

g <- ggplot(bind_rows(d2)) +
  aes(x=Day) +
  geom_line(show.legend = F) +
  coord_cartesian(xlim=c(0, max(labels$Day) + 15)) +
  ggtitle(paste0("National Trends as of ",format(end_date, "%B%e, %Y")), subtitle="Plotted individually from the first day of 5 recorded deaths") +
  xlab("Days Since First 24 Deaths") +
  facet_wrap('Country')

g + aes(y=Cases.Growth5) +
  ylab("Daily Increase in Cases") +
  scale_y_continuous(labels=percent) 

g + aes(y=Cases) + 
  ylab("Confirmed Cases")  +
  scale_y_continuous(labels=comma) 

g + aes(y=Deaths.Per100K) + 
  ylab("Deaths Per 100K Population") +
  scale_y_continuous(labels=comma) 
rm(d0,d1,d2, labels, g)
```

### EU vs USA


```{r, fig.height=3, fig.width=4}
compdata <- cvdata.i18n %>%
  filter(Country %in% c('US', 'European Union', 'Mexico', 'Sweden') &
           Date > mdy('02/26/2020')) 

compdata.last <- filter(compdata, Date == end_date)

g <- ggplot(compdata) +
  geom_line(show.legend = T) +
  xlab(NULL) +
  ylab("Daily Increase, 5 Day Average") +
  theme(legend.position = c(0.2, 0.8),
        legend.title = element_blank())

g +
   aes(y=Cases.Diff5,
      x=Date, 
      color=Country, 
      linetype=Country) +
  geom_text(data=compdata.last, 
            aes(label=paste0(Country, ': ', comma(Cases.Diff5))),
            hjust='right',
            nudge_y=2000,
            show.legend=F) +
  ylab("Growth in Confirmed Cases, 5 Day Average") +
  ggtitle("Growth in Cases",
          subtitle="Data from Johns Hopkins University Covid19 Repository")

g +
   aes(y=Cases.Diff5.Per100K,
      x=Date, 
      color=Country, 
      linetype=Country) +
  geom_text(data=compdata.last, 
            aes(label=paste0(Country, ': ', comma(Cases.Diff5.Per100K))),
            hjust='right',
            nudge_y=0,
            show.legend=F) +
  ylab("Cases per 100K") +
  ggtitle("Growth in Confirmed Cases Per 100K ",
          subtitle="Data from Johns Hopkins University Covid19 Repository")

g +
   aes(y=Deaths.Per100K,
      x=Date, 
      color=Country, 
      linetype=Country) +
  geom_text(data=compdata.last, 
            aes(label=paste0(Country, ': ', comma(Deaths.Per100K))),
            hjust='right',
            nudge_y=0,
            show.legend=F) +
  ylab("Deaths Per 100K") +
  ggtitle("Deaths Per 100K ",
          subtitle="Data from Johns Hopkins University Covid19 Repository")

g +
  aes(y=Cases,
      x=Date, 
      color=Country, 
      linetype=Country) +
  geom_text(data=compdata.last, 
            aes(label=paste0(Country, ': ', comma(Cases))),
            hjust='right',
            nudge_y=0,
            show.legend=F) +
  ylab("Confirmed Cases") +
  ggtitle("Total Confirmed Cases",
          subtitle="Data from Johns Hopkins University Covid19 Repository")

g +
  aes(y=Cases.Growth5,
      x=Date, 
      color=Country, 
      linetype=Country) +
  geom_text(data=compdata.last, 
            aes(label=paste0(Country, ': ', percent(Cases.Growth5))),
            hjust='right',
            nudge_y=0,
            show.legend=F) +
  scale_y_continuous(labels=percent) +
  ylab("Five Day Growth in Confirmed Cases") +
  ggtitle("Growth in Confirmed Cases",
          subtitle="Data from Johns Hopkins University Covid19 Repository")

```

```{r}
cvdata.i18n %>%
  filter(Country %in% c('US', 'European Union')) %>%
  ggplot() +
  aes(x=Date, color=Country) +
  geom_line(aes(y=Deaths.Diff5)) +
  xlab(NULL) +
  ylab("Daily Increase, 5 Day Average") +
  theme(legend.position = c(0.2, 0.8)) +
  ggtitle("Comparison of Deaths between EU and US",
          subtitle="Data from Johns Hopkins University Covid19 Repository")
```
```{r}
cvdata.i18n %>%
  filter(Country %in% c('US', 'European Union')) %>%
  ggplot() +
  aes(x=Date, color=Country) +
  geom_line(aes(y=Cases.Diff5.Per100K)) +
  xlab(NULL) +
  ylab("Daily Increase, 5 Day Average") +
  theme(legend.position = c(0.2, 0.8)) +
  ggtitle("Comparison of Cases per Capita Between EU and US",
          subtitle="Data from Johns Hopkins University Covid19 Repository")
```

### Analysis of Mexico
```{r}

cvdata.i18n %>%
  filter(Country == 'Mexico') %>%
  summary()
```
