---
title: "Report on United States and Covid19 Pandemic"
author: "Bill Kayser"
output:
  html_notebook: 
    fig_height: 6
    fig_width: 12
    toc: yes
  html_document: default
  pdf_document: 
    fig_height: 4
    fig_width: 7
    toc: yes
echo: TRUE

---

```{r results='hide', message=F, include=F}
library(tidyverse)
library(scales)
library(caTools)

source('../utils/labels.R')
source('../utils/utils.R')
knitr::opts_chunk$set(echo=F)
old_theme <- theme_set(ggthemes::theme_few())

```

<style type="text/css">
.table {
    width: inherit;
    max-width: 100%;
    margin-bottom: 20px;
}
.math.display {
  font-size: 28px;
}
</style>

# Load Datasets

```{r echo=T}
cvdata.us <- readRDS('../data/cvdata.us.RDS')
cvdata.us.by_state <- readRDS('../data/cvdata.us.by_state.RDS')
state.orders <- readRDS('../data/orders.events.RDS')

# Let's first set up some parameters for the graph's X axis.

start_date <- mdy('02/26/2020')
end_date <- max(cvdata.us$Date)
date.breaks <- seq(end_date,
              start_date,
              by="-1 week") %>% rev()
if (date.breaks[1] != start_date) {
  date.breaks <- c(date.breaks[1]-days(7), date.breaks)
  start_date <- date.breaks[1]
}
```

### Data Disclaimer

US Data provided by the New York Times [Covid Data Repository](https://www.nytimes.com/interactive/2020/us/coronavirus-us-cases.html).

    Note: Data are based on reports by states and counties at the time of publication. Local governments may revise reported numbers as they get new information. Some deaths may be reported by officials in two different jurisdictions. When possible, deaths have been reported here in the jurisdiction where the death occurred.
    Population and demographic data from Census Bureau.
    
# Overview of Counties

Including only counties whose population is greater than 2000.

### Top Counties in Cases per Capita


```{r}
cvdata.us %>%
  filter(!is.na(Division) & Date == end_date & Population > 2000) %>%
  arrange(desc(Cases.Per100K)) %>%
  mutate(Label=paste0(County,", ", State.Code)) %>%
  select(Label, Cases.Per100K, Deaths.Per100K, Deaths) %>%
  top_n(25) %>% knitr::kable(col.names=c('County', 'Cases Per 100K', 'Deaths Per 100K', 'Total Deaths'))

```


### Top Counties in Deaths per Capita:

```{r}
cvdata.us %>%
  filter(!is.na(Division) & Date == end_date & Population > 2000) %>%
  arrange(desc(Deaths.Per100K)) %>%
  mutate(Label=paste0(County,", ", State.Code)) %>%
  select(Label,Deaths, Deaths.Per100K) %>%
  top_n(25) %>% knitr::kable(col.names=c('County', 'Total Deaths', 'Deaths Per 100K'))
```

### Hotspots

```{r fig.height=4, fig.width=5}
hotspots <- cvdata.us %>%
  mutate(Label=paste0(County,", ", State.Code)) %>%
  filter(!is.na(Division) & Date == end_date & Population > 2000 & Cases.Diff5 >= 25) %>%
  arrange(desc(Cases.Growth5)) 

ggplot(head(hotspots, 25)) +
  aes(x=reorder(Label, Cases.Growth5), fill = gov.party, label=round(Cases.Diff5), y=Cases.Growth5) +
  geom_col(show.legend=T) +
  party_fill() +
  scale_y_continuous(position = 'right') +
  geom_text(aes(label=percent(Cases.Growth5)), size=4, nudge_y=0.01) +
  coord_flip() + 
  ylab('Average Growth 5 Day Moving Average') +
  xlab(NULL) +
  geom_text(y=0.003, color='white', hjust='inward') +
  theme(line = element_blank(), 
        panel.border = element_blank(),
        axis.text.x = element_blank(),
        legend.position=c(0.8,0.5))

```

```{r fig.height=3, fig.width=4}
cvdata.us %>%
  filter(Key %in% hotspots$Key[1:8]) %>%
  mutate(Label=reorder(as.factor(paste0(County,", ", State.Code)), Cases.Growth5)) %>%
  ggplot() +
  aes(x=Date, y=Cases, color=Label) +
  geom_line() +
  coord_cartesian(xlim=c(mdy('04-21-2020'), end_date)) +
  ggtitle("Confirmed Cases in Top 8 Hotspots") +
  xlab(NULL) +
  theme(legend.title=element_blank())


```

### Comparing Red States and Blue States

Breaking down states' data according to the party of the governor.

```{r}
cvdata.us %>%
  filter(Date == end_date & !is.na(gov.party)) %>%
  group_by(gov.party) %>%
  summarize( Count = n(),
             Cases.Diff5 = sum(Cases.Diff5),
             Cases.Per.County = round(sum(Cases) / n()),
             Growth = percent(signif(sum(Cases.Diff5) / sum(Cases), 2)),
             Cases = comma(signif(sum(Cases),3))) %>%
  select(gov.party, Count, Cases, Cases.Per.County, Growth) %>%
  knitr::kable(col.names = c("Governor's Party", "Number of Counties", "Total Cases to Date", "Cases per County", "Current 5 Day Growth"))

rm(hotspots)  
```

```{r}

cvdata.current <- cvdata.us.by_state %>%
  filter(Date == max(cvdata.us.by_state$Date)) 

model <- lm(data=cvdata.current,
   Cases.Diff5.Per100K ~ gov.party + status + Division)
summary(model)
```

```{r}

ggplot(cvdata.current) +
  aes(y=Cases.Diff5.Per100K, 
      x=Testing.Rate.Weekly, 
      size=status,
      color=gov.party) +
  scale_color_manual(values=c('R'='red', 'D'='blue')) +
  geom_point() +
  coord_flip()

ggplot(cvdata.current) +
  aes(y=Cases.Diff5, x=gov.party) +
  geom_boxplot() +
  coord_flip()
```
# State by State Comparisons

### Confirmed Cases per Capita

```{r fig.width=5, fig.height=5}
cvdata.current %>% 
  ggplot() +
  aes(x=reorder(State.Code, Cases.Per100K), y=Cases.Per100K, fill=gov.party) +
  geom_col(show.legend=F) +
  party_fill() +
  scale_y_continuous(position = 'right') +
  geom_text(aes(label=round(Cases.Per100K)), 
            hjust='right',
            size=4, 
            nudge_y=-10,
            fontface='bold',
            color='white') +
  coord_flip() + # +  scale_y_log10()
  ylab('Cases per 100K Population') +
  xlab(NULL) +
  theme(line = element_blank(), 
        panel.border = element_blank(),
        axis.text.x = element_blank(),
        legend.position=c(0.5,0.5))

```


### Change in Confirmed Cases

This plot shows the day to day change in confirmed cases.  This is important because it will tell us when we hit that peak in infections, the top of the curve that we are trying to 'flatten'.  This peak occurs when the line crosses over the X axis into negative territory.

Right now you can see New York is a long way from that but the LOESS trend line shows it starting to trend downward.


```{r echo=F, fig.height=4, fig.width=5}
topstates <- c('New York', 'Indiana', 'California', 'Oregon', 'Arizona')
g <- cvdata.us.by_state %>% 
  filter(State %in%  topstates) %>%
  mutate(State = factor(State, levels=topstates)) %>%
  arrange(State, Date) %>%
  ggplot() +
  aes(Date, Cases.Diff5, color=State, fill=State) +
  geom_line() + 
  geom_smooth(alpha=0.1, method='lm', formula= y ~ poly(x,4), fullrange=T, span=0.2, size=0.5, linetype=3) +
  ggtitle('Cases Day to Day Change in Top 5 States',
          subtitle = str_c("Reported data through ", format(max(cvdata.us$Date), "%B %d, %Y"))) +
  ylab("Daily Increase") +
  xlab(NULL)  +
  scale_y_continuous(labels = comma) +
  coord_cartesian(xlim=c(ymd('2020-03-15'), end_date)) +
  scale_x_date(breaks=date.breaks, date_labels = '%m/%d', date_minor_breaks='1 day')
g
rm(topstates)
```


```{r}
cvdata.us.by_state %>%
  filter(State %in% c('Arizona', 'Oregon', 'Indiana', 'California', 'New York')) %>%
  mutate(State = fct_drop(State)) %>%
  arrange(State, Date) %>%
  ggplot() +
  geom_line() + 
  #geom_smooth(method='loess', span=0.2, size=0.5, linetype=3) +
  ggtitle('COVID-19 Rate of Deaths Per Million, by States',
          subtitle = str_c("Reported data through ", format(max(cvdata.us$Date), "%B %d, %Y"))) +
  aes(Date, Death.Rate, color=State) +
  ylab("Daily Increase, per 1M people") +
  xlab(NULL)  +
  coord_cartesian(xlim=c(start_date, end_date)) +
  scale_x_date(breaks=date.breaks, date_labels = '%m/%d', date_minor_breaks='1 day')

```


## State Summary


```{r echo=F, fig.height=4, fig.width=5}
 state_summary_plot(cvdata.us.by_state, 
                    state = 'New York',
                    state.orders = state.orders)
 state_summary_plot(cvdata.us.by_state, 
                    state = 'Arizona',
                    state.orders = state.orders)
 state_summary_plot(cvdata.us.by_state, 
                    state = 'California',
                    state.orders = state.orders)
 state_summary_plot(cvdata.us.by_state, 
                    state = 'Texas',
                    state.orders = state.orders)
```

## Oregon


```{r echo=F, fig.height=4, fig.width=5}
 state_summary_plot(cvdata.us.by_state, 
                    state = 'Oregon',
                    state.orders = state.orders)
```

### Counties by Case Density

```{r}
cvdata.us %>%
  filter(State == 'Oregon' & Date == end_date) %>%
  arrange(desc(Cases.Per100K)) %>%
  select(County, Cases, Deaths, Cases.Per100K, Population) %>%
  knitr::kable(col.names = c('County', 'Cases', 'Deaths', 'Cases Per 100K', 'Total Population'), format.args = c(big.mark=',', zero.print=''))

```

### Counties by Recent Growth
```{r}
cvdata.us %>%
  filter(State == 'Oregon' & Date == end_date & Cases.Diff5 > 0) %>%
  arrange(desc(Cases.Diff5.Per100K)) %>%
  mutate(Cases.Diff5.Per100K = round(Cases.Diff5.Per100K,1)) %>%
  select(County, Cases, Deaths, Cases.Diff5, Cases.Diff5.Per100K, Population) %>%
  knitr::kable(col.names = c('County', 'Cases', 'Deaths', '5 Day Increase', 'Per Capita Increase', 'Population'), format.args = c(big.mark=',', zero.print=''))
```


## State by State Confirmed Cases

```{r, fig.width=6, fig.height=6}
today <- filter(cvdata.us) %>%
  filter(Date == end_date & !is.na(Region) & Population > 500)

averages.by_state <- tapply(today$Cases.Per100K, today$State.Code, median) %>% sort() 
ggplot(today) +
  aes(x = fct_relevel(State.Code, names(averages.by_state)), y = Cases.Per100K, fill = gov.party) +
  geom_boxplot() +
  coord_flip(ylim=c(0,2000)) +
    party_fill() +
  ggtitle("") +
  ggtitle('Cases Per Capita for Counties, State by State')

rm(today, averages.by_state)
```

### Infections Per Capita

Let's chart the current confirmed cases and deaths as a percentage of population.

```{r}
cvdata.us.by_state %>%
  filter(Date == end_date) %>%
  mutate(Infections.Per.100_000 = round(100000 * Cases / Population),
         Deaths.Per.100_000 = round(100000 * Deaths / Population, 1)) %>%
  select(State, Infections.Per.100_000, Deaths.Per.100_000) %>%
  arrange(desc(Infections.Per.100_000)) %>%
  head(10)

```

### Infection Rate by Region

```{r fig.width=5}
cvdata.us.by_state %>% 
  filter(Date == end_date) %>%
  ggplot() +
  aes(x=Region, y=Deaths.Per100K, fill=Region) +
  geom_boxplot() +
#  scale_y_log10(labels=comma) +
#  scale_x_discrete(labels=NULL) +
  xlab('Region') +
  ggtitle('Deaths Per 100K, by Region')
```

### Infection Rate by Governor's Party

```{r fig.width=5}
cvdata.us.by_state %>% 
  filter(Date == end_date) %>%
  ggplot() +
  aes(x=gov.party, y=Deaths.Per100K, fill=gov.party) +
  party_fill() +
  geom_boxplot(show.legend=F) +
  xlab('Party') +
  ggtitle('Deaths Per 100K, by Party')
```
### Current Infection Growth Rate by Governor's Party

```{r fig.width=5}
cvdata.us.by_state %>% 
  filter(Date == end_date) %>%
  ggplot() + 
  aes(x=gov.party, y=Cases.Growth5, fill=gov.party) +
  party_fill() +
  geom_boxplot(show.legend = F) +
  scale_y_continuous(labels=percent) +
  xlab('Party') +
  ggtitle('Growth in Cases, by Party')
```
### Scatter Plots

```{r, fig.width=5, fig.height=8}

lut <- cvdata.us.by_state %>%
  group_by(gov.party, Division) %>%
  summarize(States = list(unique(State)))

cvdata.us.by_state %>%
  filter(State != 'District of Columbia') %>%
  ggplot() +
  aes(x=Cases.Per100K, y=Cases.Diff5.Per100K, color=State) +
  geom_line(show.legend = F) +
  facet_grid(Division ~ gov.party)
  
```

### Report on Lake Placid

Essex and Franklin counties seem to be closest.

```{r, fig.height=5}
our_counties <- filter(cvdata.us, 
                     (State == "New York" & County %in% c('Essex', 'Franklin', 'Clinton')) |
                     (State == "Indiana" & County %in% c('Hamilton')) |                       
                       (State == "California" & County %in% c('Contra Costa')) |
                     (State == "Arizona" & County %in% c('Maricopa')) |
                       (State == "Oregon" & County %in% c('Clackamas', 'Multnomah', 'Washington')))
labels <- filter(our_counties, Date == end_date)

g <- ggplot(our_counties) +
  aes(x=Date, y=Cases.Per100K, color=County) +
  xlab(NULL)  +
  coord_cartesian(xlim=c(mdy('03/15/2020'), end_date)) +
  ggtitle('Kayser Family Counties',
          subtitle = str_c("Reported data through ", format(max(cvdata.us$Date), "%B %d, %Y"))) +
  geom_line(aes(linetype=State)) + 
  geom_label(data=labels, aes(label=Key), hjust='right', show.legend=F, nudge_x = -1) +
  scale_x_date(breaks=date.breaks, date_labels = '%m/%d', date_minor_breaks='1 day')

g +
  aes(y = Cases.Per100K) +
  ylab("Confirmed Cases per 100K Population")

g +
  aes(y = Cases) +
  ylab("Confirmed Cases")

rm(labels)
```




```{r}

LP.Cases.Per100K <- filter(our_counties, Date == end_date) %>% 
  group_by(State) %>%
  summarize(Deaths = sum(Deaths),
            Cases = sum(Cases),
            Population = sum(Population)) %>%
  ungroup() %>%
  mutate(Cases.Per100K = Cases * 100000 / Population)

rm(our_counties)
```

Total deaths to date in all three counties: `r LP.Cases.Per100K$Deaths`.

Total cases per 100K people: `r round(LP.Cases.Per100K$Cases.Per100K, 1)`

## Population Density vs. Infection Rate

```{r fig.width=5}
filter(cvdata.us, Date == end_date & !is.na(Population.Density)) %>%
  mutate(Deaths.Per100K = Deaths * 100000 / Population) %>%
  filter(Deaths.Per100K > 0) %>%
ggplot() +
  aes(x=Population.Density, y=Cases.Per100K, show.legend=F) +
  geom_point(size=0.2) +
  geom_smooth(method='lm', formula = y ~ x, color='red', size=0.5) +
  scale_y_log10(labels=comma) + 
  scale_x_log10() +
  ggtitle(paste0('Confirmed Cases vs. Population Density, as of ', format(end_date, "%D")),
          subtitle='Data in Log Scale') +
  xlab('Population per Mile') +
  ylab('Cases per 100K population')
```


```{r fig.width=5, fig.height=4}

testdata.ts <- cvdata.us.by_state %>%
  mutate(Test.Positive.Rate = positive/(positive + negative),
         Tests = positive + negative,
         Tests.Per.Capita = 100000 * Tests/Population) %>%
  arrange(desc(Test.Positive.Rate)) 

testdata <- testdata.ts %>%
  filter(Date==max(cvdata.us.by_state$Date)) %>%
  mutate(State = fct_reorder(State, Tests.Per.Capita))  


ggplot(testdata) +
  aes(x=State.Code, y=Tests.Per.Capita, fill=gov.party) +
  geom_bar(stat='identity') +
  party_fill() +
  scale_y_continuous(labels=comma, position='right') +
  ggtitle('Tests conducted per 100K population') +
  coord_flip() 

testdata$State = fct_reorder(testdata$State, testdata$Test.Positive.Rate)

ggplot(testdata) +
  aes(x=State, y=Testing.Positive.Rate, fill=gov.party) +
  party_fill() +
  geom_bar(stat='identity') +
  scale_y_continuous(labels=percent, position='right') +
  coord_flip() + #ylim=c(0,0.25))  +
  theme(line = element_blank(), 
        panel.border = element_blank(),
#        axis.text.x = element_blank(),
        legend.position=c(0.75,0.4))

```

```{r, fig.width=4, fig.height=3.5}

window_width <- 4
timerange <- 60
history <- filter(testdata.ts, 
                  !is.na(total.Diff5) &                  # Filter missing values
                    total.Diff5 > 0 &                    # Filter missing values
                    Date > today() - days(timerange))    # Recent data only

history.1 <- split(history, history$State) %>% 
  lapply(function(df) {
    arrange(df, Date) %>%
      mutate(positiveRate = positive.Diff5 / total.Diff5)
  }) %>% 
  bind_rows() %>%
  filter(positiveRate < 0.55)  # Remove outliers

filter(history.1, State %in% c('New York', 'New Jersey', 'Massachusetts', 'Arizona', 'Texas', 'Georgia', 'South Carolina', 'Florida')) %>%
  ggplot() +
  aes(x=Tests.Per.Capita, y=positiveRate, color=State) +
  ggtitle(subtitle = paste0(window_width, ' day moving average over last ', timerange, ' days'),
          paste0('Per Capita Testing vs. Test Positive Rate')) +
  xlab('Tests per 100K Population') +
  ylab('Rate of Positive Tests') +
  geom_line(show.legend = F, size=0.2) +
  geom_point(show.legend = F, size=0.6) +
  scale_y_continuous(labels=percent) +
  geom_label(data=~group_by(.x, State.Code) %>% summarize(State = last(State),
                                                          Date = last(Date),
                                                          positiveRate=last(positiveRate),
                                                          Tests.Per.Capita=last(Tests.Per.Capita)),
             aes(label=State.Code), show.legend=F)  

filter(history.1, State %in% c('Oregon', 'New York', 'New Jersey', 'Massachusetts')) %>%
  ggplot() +
  aes(x=Tests.Per.Capita, y=positiveRate, color=State) +
  ggtitle(subtitle = paste0(window_width, ' day moving average over last ', timerange, ' days'),
          paste0('Per Capita Testing vs. Test Positive Rate')) +
  xlab('Tests per 100K Population') +
  ylab('Rate of Positive Tests') +
  geom_line(show.legend = F, size=0.2) +
  geom_point(show.legend = F, size=0.6) +
  scale_y_continuous(labels=percent) +
  geom_label(data=~group_by(.x, State.Code) %>% summarize(State = last(State),
                                                          Date = last(Date),
                                                          positiveRate=last(positiveRate),
                                                          Tests.Per.Capita=last(Tests.Per.Capita)),
             aes(label=State.Code), show.legend=F)  


```

```{r, fig.width=5, fig.height=4}

df <- filter(history.1, State %in% c('Indiana', 'California', 'Arizona', 'Oregon', 'New York')) 
#filter(history.1, State %in% c('California', 'Maine', 'New Jersey')) %>% 
#  select(State, Date, contains('Increase'), contains('Rate')) %>% view()

g <- split(history.1, history.1$Region) %>% 
  lapply(function(df) {
    df.current <- filter(df, Date == max(df$Date))
    ggplot(df) +
      aes(x=Tests.Per.Capita, y=positiveRate, color=State) +
      ggtitle(subtitle = paste0(window_width, ' day moving average over last ', timerange, ' days'),
              paste0('Per Capita Testing vs. Test Positive Rate for States in ', df$Region[1], ' Region')) +
      xlab('Tests per 100K Population') +
      ylab('Rate of Positive Tests') +
      geom_line(show.legend = F, size=0.2) +
      geom_point(show.legend = F, size=0.6) +
      scale_y_continuous(labels=percent) +
      geom_label(data=df.current, aes(label=State.Code), show.legend=F) #+
      #coord_cartesian(xlim=c(0, max(testdata$Tests.Per.Capita))) 
  }) 

g
rm(history, history.1, timerange, window_width, df)
```

## Arizona Counties

```{r}
filter(cvdata.us, State == 'Arizona' & Cases.Per100K > 0.1) %>%
  ggplot() +
  aes(
    x = Date,
    y = Cases.Per100K,
    color = County,
    linetype = County
  ) +
  geom_line(show.legend = F) +
  coord_cartesian(xlim = c(mdy('03-24-2020'), Sys.Date())) +
  geom_text(
    data = ~ filter(.x, Date == end_date),
    aes(x = Date, y = Cases.Per100K, label = County),
    show.legend = F,
    nudge_y = 8
  )

```

# Cases and Recoveries

```{r, fig.width=4}

filter(cvdata.us.by_state, State %in% head(arrange(testdata, desc(Cases))$State, 6)) %>%
  ggplot() +
  aes(x=Date) +
  geom_line(aes(y=Cases)) +
  geom_line(aes(y=Est.Recovered), linetype=2) +
  geom_line(aes(y=Recovered), linetype=4) +
  theme(legend.title = element_blank(),
        legend.position = c(0.2, 0.8)) +
  coord_cartesian(xlim=c(mdy('03-15-2020'), today())) +
  scale_y_continuous(label=comma) +
  ylab('Cases') + xlab(NULL) +
  theme_light() +
  facet_wrap('State', 2)
rm(testdata, testdata.ts)         
```

# Bubble Charts

Plot the number of cases outside the number of recoveries.

```{r fig.width=2, fig.height=2}

df <- filter(cvdata.us.by_state, Date == max(Date)) %>%
  arrange(desc(Long)) %>%
  mutate(col = ((row_number()-1) %% 9), 
         row = ((row_number()-1) %/% 9),
         label_size = 0.05* pmax(Cases/Population, 0.002))
         
df$State.Code = fct_inorder(df$State.Code)

g <- ggplot(df) +
  aes(x=col, y=5-row) +
  geom_point(aes(size = Cases/Population), color='red', alpha=0.5, show.legend=F) +
  xlab(NULL) + ylab(NULL) +
  coord_cartesian(xlim=c(-0.5,8.5), ylim=c(-0.5,5.5)) +
  scale_size(range=c(0, 20), limits=c(0,max(cvdata.us.by_state$Cases/cvdata.us.by_state$Population))) +
  theme(panel.grid = element_blank(),
        panel.grid.major = element_blank(),
        axis.text = element_blank(),
        legend.title = element_blank()) 

g + 
  geom_point(aes(size = Recovered/Population), color='green', alpha=0.5, show.legend=F) +
  geom_text(aes(label=State.Code, size=label_size), nudge_x=-0.0, show.legend = F) +
  ggtitle('Cases and Reported Recoveries per Capita ', 'Recoveries only reported from some states')

g + 
  geom_point(aes(size = Est.Recovered/Population), color='green', alpha=0.5, show.legend=F) +
  geom_text(aes(label=State.Code, size=label_size), nudge_x=-0.0, show.legend = F) +
  ggtitle('Cases and Estimated Recoveries per Capita', 'Recoveries estimated with 26 day median recovery time')

rm(df)

```

